---
title: "Analysis of Thomas Poulton’s Descriptions"
author: "Bodo Winter"
date: "3/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preprocessing

Load data and packages:

```{r, message = FALSE, warning = FALSE}
# Packages:

library(tidyverse)
library(tidytext)
library(lavaan)
library(effsize)
library(corrplot)
library(gridExtra)

# Data:

scents <- read_csv('../data/by_scent_cleaned.csv')
war <- read_csv('../data/warriner_2013_cleaned.csv')
lyn <- read_csv('../data/lynott_norms.csv')
```

## Confirmatory factor analysis of codability:

Z-score variables that enter the CFA:

```{r}
scents <- mutate(scents,
                 Length_z = scale(Length),
                 TextEntropy_z = scale(TextEntropy),
                 Jaccard_z = scale(jaccard))
```


Create formula for latent variable 'codability' that is indicated by the three measures (length, text entropy, and inter-speaker agreement measured by jaccard similarity).

```{r}
CFA_formula <- 'Codability =~ Length_z + TextEntropy_z + Jaccard_z'
```

Fit this CFA:

```{r}
CFA_fit <- cfa(CFA_formula, data = scents)
CFA_fit
```

Summarize it:

```{r}
summary(CFA_fit, fit.measures = TRUE, standardized = TRUE)
```

## Structural equation model with haptic:

Haptic SEM formula, regressing haptic ratings per description onto the latent factor codability:

```{r}
haptic_SEM_formula <- '
  # measurement model:
  Codability =~ Length_z + TextEntropy_z + Jaccard_z

  # regression:
  Haptic ~ Codability'
```

Fit SEM:

```{r}
haptic_SEM_fit <- cfa(haptic_SEM_formula, data = scents)
```

Summarize:

```{r}
summary(haptic_SEM_fit)
```

R-squared:

```{r}
inspect(haptic_SEM_fit, 'r2')
```

## Structural equation model with source-based:

Scale source:

```{r}
scents <- mutate(scents,
                 Source_z = Source - mean(Source),
                 Source_z = Source_z / sd(Source_z))
```

Source SEM formula, regressing haptic ratings per description onto the latent factor codability:

```{r}
source_SEM_formula <- '
  # measurement model:
  Codability =~ Length_z + TextEntropy_z + Jaccard_z

  # regression:
  Source_z ~ Codability'
```

Fit SEM:

```{r}
source_SEM_fit <- cfa(source_SEM_formula, data = scents)
```

Summarize:

```{r}
summary(source_SEM_fit)
```

Need to investigate warning messages.

Check AICs of the CFA versus the two SEMs:

```{r}
AIC(CFA_fit)
AIC(source_SEM_fit)
AIC(haptic_SEM_fit)
```

Is this actually interpretable? (not sure what counts as the metric in an SEM context)

Double check:

```{r}
CFA_scores <- predict(CFA_fit)[, 1]
summary(lm(scents$Source ~ CFA_scores))
summary(lm(scents$Haptic ~ CFA_scores))
```

## Analysis of efficient codability measures:

First, extract those:

```{r}
codability <- select(scents, jaccard, TextEntropy, Length)
```

Correlate these with each other:

```{r}
corrs <- cor(codability)
```

What are the correlations?

```{r}
round(corrs, 2)
```

Extract a conjoined measure. First, z-score:

```{r}
codability_z <- apply(codability, 2, FUN = scale)
```

Perform a PCA on this:

```{r}
cod_PCA <- prcomp(codability_z, center = TRUE, scale = TRUE)
```

Check variance explained:

```{r}
summary(cod_PCA)
```

Check loadings:

```{r}
round(cod_PCA$rotation, 2)
```

Get the first component:

```{r}
scents$Codability <- cod_PCA$x[, 1]
```


## Correlate codability with the rest:

Check codability correlations with the different sensory modalities:

```{r}
with(scents, cor.test(Codability, Haptic))
with(scents, cor.test(Codability, Olfactory))
with(scents, cor.test(Codability, Gustatory))
with(scents, cor.test(Codability, Visual))
with(scents, cor.test(Codability, Auditory))
```

Do haptic and exclusivity correlate?

```{r}
with(scents, cor.test(Haptic, Excl_nochem))
```

Check correlations with source:

```{r}
with(scents, cor.test(Codability, Source))
with(scents, cor.test(Codability, Excl_nochem))
```

What about Source and haptic?

```{r}
with(scents, cor.test(Haptic, Source))
with(scents, cor.test(Source, Abstract))
with(scents, cor.test(Haptic, Abstract))
```


## Against baseline comparisons:

Exclusivity against with baseline:

```{r}
chem <- c('Gustatory', 'Olfactory')

mean(scents$Excl_nochem, na.rm = TRUE)
mean(filter(lyn, !DominantModality %in% chem)$ModalityExclusivity)

t.test(scents$Excl_nochem, filter(lyn, !DominantModality %in% chem)$ModalityExclusivity, var.equal = TRUE)
```

Effect size:

```{r}
cohen.d(scents$Excl_nochem, filter(lyn, !DominantModality %in% chem)$ModalityExclusivity, paired = FALSE, na.rm = TRUE)
```

Absolute valence against baseline:

```{r}
mean(scents$AbsVal, na.rm = TRUE)
mean(abs(war$Val_z))

t.test(scents$AbsVal, abs(war$Val_z), var.equal = TRUE)
```

Effect size:

```{r}
cohen.d(scents$AbsVal, abs(war$Val_z), paired = FALSE, na.rm = TRUE)
```

Valence against baseline:

```{r}
mean(scents$Val_z, na.rm = TRUE)
mean(war$Val_z)

t.test(scents$Val_z, mu = mean(war$Val_z))
```

## Make plots of these distributions:

Put both together for comparison, absolute valence:

```{r}
# Extract valence only:

absval_only <- select(scents, AbsVal)
absval_only$Type <- 'Smell descriptions'

# Extract valence from warriner data frame:

absval_baseline <- war %>%
  mutate(AbsVal = abs(Val_z)) %>% 
  select(AbsVal)
absval_baseline$Type <- 'Warriner et al. (2013)'

# Merge:

absval <- bind_rows(absval_only, absval_baseline)
```

Make a plot of this:

```{r, fig.width = 10, fig.height = 6}
absval %>% ggplot(aes(x = AbsVal, fill = Type)) +
  geom_density(alpha = 0.5) + scale_fill_viridis_d() +
  theme_minimal() +
  xlab('Absolute valence') +
  ylab('Density') +
  theme_minimal() + 
  theme(axis.text.x = element_text(face = 'bold', size = 12),
        axis.text.y = element_text(face = 'bold', size = 12),
        axis.title.x = element_text(face = 'bold', size = 16,
                                    margin = margin(t = 20, r = 0, b = 0, l = 0)),
        axis.title.y = element_text(face = 'bold', size = 16,
                                    margin = margin(t = 0, r = 20, b = 0, l = 0)),
        plot.title = element_text(face = 'bold', size = 22),
        legend.title = element_blank())  

ggsave('../figures/absolute_valence.pdf', width = 9, height = 6)
```

Put both together for comparison, modality exclusivity:

```{r}
# Extract valence only:

excl_only <- select(scents, Excl_nochem)
excl_only$Type <- 'Smell descriptions'
excl_only <- excl_only %>% rename(Excl = Excl_nochem)

# Extract valence from warriner data frame:

excl_baseline <- lyn %>%
  filter(!(DominantModality %in% chem)) %>% 
  rename(Excl = ModalityExclusivity) %>% 
  select(Excl)
excl_baseline$Type <- 'Lynott & Connell (2009)'

# Merge:

excl <- bind_rows(excl_only, excl_baseline)
```

Make a plot of this:

```{r, fig.width = 10, fig.height = 6}
excl %>% ggplot(aes(x = Excl, fill = Type)) +
  geom_density(alpha = 0.5) +
  scale_fill_viridis_d() +
  xlab('Modality exclusivity') +
  ylab('Density') +
  theme_minimal() + 
  theme(axis.text.x = element_text(face = 'bold', size = 12),
        axis.text.y = element_text(face = 'bold', size = 12),
        axis.title.x = element_text(face = 'bold', size = 16,
                                    margin = margin(t = 20, r = 0, b = 0, l = 0)),
        axis.title.y = element_text(face = 'bold', size = 16,
                                    margin = margin(t = 0, r = 20, b = 0, l = 0)),
        plot.title = element_text(face = 'bold', size = 22),
        legend.title = element_blank())  
ggsave('../figures/exclusivity_valence.pdf', width = 9, height = 6)
```



