---
title: "Analysis of Thomas Poulton’s Descriptions"
author: "Bodo Winter"
date: "3/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preprocessing

Load data and packages:

```{r, message = FALSE, warning = FALSE}
# Packages:

library(tidyverse)
library(tidytext)
library(vegan) # for entropy
library(textreuse) # for document similarity

# Data:

smell <- read_csv('../data/02_coded-descriptions-IRR.csv')
war <- read_csv('../data/warriner_2013_cleaned.csv')
lyn <- read_csv('../data/lynott_norms.csv')
```

Clean column names:

```{r}
colnames(smell) <- str_to_lower(colnames(smell))
colnames(smell)[3] <- 'ID'
```

Tokenize descriptions:

```{r}
tokens <- smell %>% unnest_tokens(Word, text)
```

Description length:

```{r}
lengths <- tokens %>% count(ID)
```

Merge this back into there:

```{r}
smell <- lengths %>%
  rename(Length = n) %>%
  right_join(smell)
```

How many smell descriptions had source-based language?

```{r}
sum(smell$source) # total sum
sum(ifelse(smell$source >= 1, 1, 0))
nrow(smell)
sum(ifelse(smell$source >= 1, 1, 0)) / nrow(smell)
```


Merge this with Lynott & Connell (2009) norms:

```{r}
# Clean:

lyn <- lyn %>%
  select(Property, VisualStrengthMean:GustatoryStrengthMean,
         DominantModality, ModalityExclusivity) %>% 
  rename(Word = Property, Excl = ModalityExclusivity, Mod = DominantModality)

# Simplify perceptual strength column names:

colnames(lyn) <- str_replace(colnames(lyn), 'StrengthMean', '')

# Join:

tokens <- left_join(tokens, lyn)
```

Create absolute valence measure in the Warriner et al. (2013) data:

```{r}
war <- mutate(war, AbsVal = abs(Val_z))
```

Merge with valence:

```{r}
tokens <- left_join(tokens, war)
```

How many of those tokens are sensory adjectives from Lynott & Connell (2009)?

```{r}
sum(!is.na(tokens$Mod))
sum(!is.na(tokens$Mod)) / nrow(tokens)
```

How many descriptions have at least one of the adjectives?

```{r}
filter(tokens, !is.na(Mod)) %>% count(ID) %>% nrow()
length(unique(tokens$ID))

# Proportion:
filter(tokens, !is.na(Mod)) %>% count(ID) %>% nrow() / length(unique(tokens$ID))
```

For how many of the tokens do we have valence information?

```{r}
nrow(tokens)
sum(!is.na(tokens$AbsVal))
sum(!is.na(tokens$AbsVal)) / nrow(tokens)
```

How many descriptions have at least one of words with valence information?

```{r}
filter(tokens, !is.na(AbsVal)) %>% count(ID) %>% nrow()
length(unique(tokens$ID))

# Proportion:
filter(tokens, !is.na(AbsVal)) %>% count(ID) %>% nrow() / length(unique(tokens$ID))
```

Calculate average valence and absolute per description:

```{r}
smell <- tokens %>% group_by(ID) %>% 
  summarize(Val_z = mean(Val_z, na.rm = TRUE),
            AbsVal = mean(AbsVal, na.rm = TRUE),
            Visual = mean(Visual, na.rm = TRUE),
            Haptic = mean(Haptic, na.rm = TRUE),
            Auditory = mean(Auditory, na.rm = TRUE),
            Gustatory = mean(Gustatory, na.rm = TRUE),
            Olfactory = mean(Olfactory, na.rm = TRUE)) %>% right_join(smell)
```

Multisensory without taste and smell words measure:

```{r}
tokens_nochem <- tokens %>% filter(!(Mod %in% c('Gustatory', 'Olfactory')))
```

Calculate exclusivity:

```{r}
smell <- tokens_nochem %>% group_by(ID) %>% 
  summarize(Excl_nochem = mean(Excl, na.rm = TRUE)) %>% right_join(smell)
```

## Calculate entropy:

Get rid of stop words:

```{r}
no_stops <- tokens %>% anti_join(stop_words, by = c('Word' = 'word'))
```

Spread the words:

```{r}
no_stops <- no_stops %>% select(scent, Word)

scent_word <- no_stops %>% count(scent, Word) %>% spread(key = 'Word', value = 'n', fill = 0)
```

Calculate entropy:

```{r}
entropies <- diversity(scent_word, index = 'shannon', MARGIN = 1)
scents <- arrange(smell, scent) %>% pull(scent) %>% unique()
scents <- tibble(scent = scents, TextEntropy = entropies)
```

## Calculate Jaccard similarity across descriptions:

Get Jaccard similarity for each element:

```{r}
tokens_nostop <- anti_join(tokens, stop_words, by = c('Word' = 'word'))
```

Compute Jaccard similarity across descriptions for each smell:

```{r}
all_smells <- unique(tokens_nostop$scent)
smell_avg <- numeric(length(all_smells))
for (i in seq_along(all_smells)) {
  this_smell <- filter(tokens_nostop, scent == all_smells[i])
  
  # Extract IDs from this:
  
  all_IDS <- unique(this_smell$ID)
  
  # Create a similarity matrix for the number of IDS:
  
  M <- matrix(numeric(length(all_IDS) * length(all_IDS)), nrow = length(all_IDS))
  
  for (j in seq_along(all_IDS)) {
    this_ID <- all_IDS[j]
    this_words <- filter(this_smell, ID == this_ID)$Word
    # other_IDs <- all_IDS[all_IDS != this_ID]
    for (k in seq_along(all_IDS)) {
      comparison_words <- filter(this_smell, ID == all_IDS[k])$Word
      M[j, k] <- jaccard_similarity(this_words, comparison_words)
    }
  }
  # Compute average across entire matrix (without diagonal, which is always 1):
  
  diag(M) <- NA
  smell_avg[i] <- mean(M, na.rm = TRUE)
}
```

Put those into a table:

```{r}
scent_sim <- tibble(scent = all_smells, jaccard = smell_avg)
```

Join with other measures:

```{r}
scents <- left_join(scent_sim, scents)
```

## Create averages table:

Averages:

```{r}
scents <- smell %>%
  group_by(scent) %>% 
  summarize(Length = mean(Length, na.rm = TRUE),
    Val_z = mean(Val_z, na.rm = TRUE),
    AbsVal = mean(AbsVal, na.rm = TRUE),
    Visual = mean(Visual, na.rm = TRUE),
    Haptic = mean(Haptic, na.rm = TRUE),
    Auditory = mean(Auditory, na.rm = TRUE),
    Gustatory = mean(Gustatory, na.rm = TRUE),
    Olfactory = mean(Olfactory, na.rm = TRUE),
    Source = mean(source),
    Abstract = mean(abstract),
    Excl_nochem = mean(Excl_nochem, na.rm = TRUE)) %>%
  right_join(scents)
```

Outsource table:

```{r}
write_csv(scents, '../data/by_scent_cleaned.csv')
```




