---
title: "Analysis of Thomas Poulton’s Descriptions"
author: "Bodo Winter"
date: "3/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analysis of sensory words:

Load data and packages:

```{r, message = FALSE, warning = FALSE}
# Packages:

library(MASS) # for negative binomial model
library(tidyverse)
library(tidytext)
library(lme4)

# Data:

smell <- read_csv('../data/02_coded-descriptions-IRR.csv')
lyn <- read_csv('../data/lynott_norms.csv')
```

Clean column names:

```{r}
colnames(smell) <- str_to_lower(colnames(smell))
colnames(smell)[3] <- 'ID'
```

Tokenize descriptions:

```{r}
tokens <- smell %>% unnest_tokens(Word, text)
```

Description length:

```{r}
lengths <- tokens %>% count(ID)
```

Merge this with Lynott & Connell (2009) norms:

```{r}
# Clean:

lyn <- lyn %>%
  dplyr::select(Property, VisualStrengthMean:GustatoryStrengthMean,
         DominantModality, ModalityExclusivity) %>% 
  rename(Word = Property, Excl = ModalityExclusivity, Mod = DominantModality)

# Simplify perceptual strength column names:

colnames(lyn) <- str_replace(colnames(lyn), 'StrengthMean', '')

# Join:

tokens <- left_join(tokens, lyn)
```

Add number of sensory adjectives per each scent:

```{r}
scent_modality <- tokens %>% 
  count(scent, Mod) %>%
  filter(!is.na(Mod)) %>% 
  spread(value = 'n', key = 'Mod', fill = 0)
```

How many scents have X modality been used to describe it?

```{r}
# Counts:

mod_counts <- apply(scent_modality[, -1], 2, FUN = function(x) sum(x != 0))
Auditory <- 0
names(Auditory) <- 'Auditory'
mod_counts <- c(mod_counts, Auditory)
mod_counts

# Percentages:

str_c(round(mod_counts / 42, 2) * 100, '%')
```

Chi-square test of distribution of sensory words:

```{r}
chisq.test(mod_counts)
```

How many sensory words are used on average per sense:

```{r}
round(colMeans(scent_modality[, -1]), 1)
```

Create a negative binomial model for this:

```{r}
sense_counts <- tokens %>% 
  count(scent, Mod) %>%
  filter(!is.na(Mod))
```

Unequal distribution:

```{r}
pois_mdl <- glmer(n ~ Mod + (1|scent), data = sense_counts, family = 'poisson')
pois_null <- glmer(n ~ 1 + (1|scent), data = sense_counts, family = 'poisson')
anova(pois_null, pois_mdl)
```

## What are the adjectives?

Extract unique types per sensory modality:

```{r}
# Gustatory words:

gus_words <- sort(unique(filter(tokens, Mod == 'Gustatory')$Word))

# Olfactory words:

olf_words <- sort(unique(filter(tokens, Mod == 'Olfactory')$Word))

# Haptic words:

hap_words <- sort(unique(filter(tokens, Mod == 'Haptic')$Word))

# Visual words:

vis_words <- sort(unique(filter(tokens, Mod == 'Visual')$Word))
```

Print:

```{r}
gus_words
olf_words
hap_words
vis_words
```

Compare these visual words to the baseline visual words in Lynott & Connell (2009):

```{r}
mean(filter(lyn, Word %in% vis_words)$Excl)
mean(filter(lyn, Mod == 'Visual')$Excl)

t.test(filter(lyn, Word %in% vis_words)$Excl, filter(lyn, Mod == 'Visual', !(Word %in% vis_words))$Excl,
       var.equal = TRUE)
```






